Estimating initial state and quark-gluon plasma medium properties \\ using a hybrid model with nucleon substructure \\ calibrated to p-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV

### 2018-07-23 paper/substructure.tex
More work on the results section

\newcommand{\trento}{\textsc{T\raisebox{-0.5ex}{R}ENTo}}
\newcommand{\sqrts}{\sqrt{s_{NN}}}
\newcommand{\nch}{N_\text{ch}}
\newcommand{\ntrk}{N_\text{trk}^\text{offline}}
\newcommand{\sigmaf}{\sigma_\mathrm{fluct}}
\newcommand{\X}{\chi_\mathrm{struct}}
\newcommand{\taufs}{\tau_\mathrm{fs}}
\newcommand{\dmin}{d_\mathrm{min}}
\newcommand{\T}{\tilde{T}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\vnk}[2]{v_#1\{#2\}}
\newcommand{\paddedhline}{\noalign{\smallskip}\hline\noalign{\smallskip}}
\newcommand{\order}[1]{$\mathcal O(10^{#1})$}
\newcommand{\note}{\textcolor{theblue}{[?]}}

@@

\begin{abstract}
We posit a unified hydrodynamic and microscopic description of the quark-gluon plasma (QGP) produced in ultrarelativistic p-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference. Specifically, we model the dynamics of both collision systems using initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$ and flow cumulants.
We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that the flow produced develops at length scales smaller than a proton.
Posterior estimates for the model's input parameters are obtained, and new insights into the temperature dependence of the QGP transport coefficients and event-by-event structure of the proton are discussed.
\end{abstract}

@@

Ultrarelativistic nuclear collisions between one light-ion and one heavy-ion, e.g.\ $^3$He-Au and p-Pb collisions, generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions where collectivity is commonly explained by the existence of hydrodynamic flow \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla, Adare:2015ctn}.
This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP) \cite{Bozek:2011if, Bozek:2013uha}, and that flow might develop at length scales smaller than a single proton \note.

@@

Independently, studies of coherent and incoherent $J/\psi$ production based on a color dipole picture of vector meson production were used to simultaneously constrain both the average color charge density of the proton as well as its event-by-event fluctuations in a saturation based framework \cite{Mantysaari:2016ykx, Aaron:2009aa, Abramowicz:2015mha}.
Initial condition studies using the IP-Glasma model of color-glass condensate effective field theory \cite{Schenke:2012wb} simultaneously demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
In addition, it was recently shown that hydrodynamic simulations using IP-Glasma initial conditions with color charge fluctuations calibrated to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha} provide a good description of collectivity in small and large collision systems \cite{Schenke:2018qmc}.

@@

With these considerations in mind, we present progress towards a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free-streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$ and anisotropic flow coefficients of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.

@@

Assuming our two protons collide at the sampled impact parameter $\mathbf{b}$, we assign each a \emph{fluctuated} thickness
\begin{equation}
  \label{eq:fluctuated_thick}
  \T_{A, B}(\x) \equiv \int dz\, \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c \,(\mathbf{x} - \mathbf{x_i} \pm \mathbf{b}/2),
\end{equation}
equal to the beam-integrated proton density in Eqn.~\eqref{eq:rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their physical origin is not yet well understood.

Within the eikonal approximation, the initial entropy deposited at midrapidity at proper time $\tau=0^+$~fm/$c$ can be represented by some mapping ${f(\T_A, \T_B) \mapsto dS/dy}$.
A natural first guess for the function $f$ is an arithmetic mean $f \propto (\T_A + \T_B)/2$ of participant nuclear matter $\T_{A,B}$, which is simply a wounded nucleon model up to meaningless factor of two in the overall normalization.
The wounded nucleon model was in fact one of the first such mappings used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thickness $\T_{A,B}$ \cite{Kharzeev:2000ph}.

A simple remedy is to replace the arithmetic mean in the wounded nucleon model with a more flexible parametrization
\begin{equation}
  \label{eq:gmean}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto \left( \frac{\T_A^p + \T_B^p}{2} \right)^{1/p},
\end{equation}
based on a family of functions known as the generalized mean(s).
This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric and harmonic means:
\begin{equation}
  \newlength{\extraspace}
  \setlength{\extraspace}{0.5ex}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto
  \begin{cases}
    \max(T_A, T_B) & p \rightarrow +\infty, \\[\extraspace]
    (T_A + T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
    \sqrt{T_A T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
    2\, T_A T_B/(T_A + T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
    \min(T_A, T_B) & p \rightarrow -\infty.
  \end{cases}
\end{equation}
Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
Of course the model is \emph{not} a substitute for first principle theory calculations, and it may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.

@@

\begin{figure}[b]
  \includegraphics{coupling}
  \caption{Cartoon of the free-streaming approximation for hydrodynamic initialization. The initial state is free-streamed for proper time $\taufs$ (zero-coupling) before it is matched to hydrodynamics (strong-coupling). This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
  \label{fig:coupling}
\end{figure}

@@

For the prior $P(H_\x)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Tbl.~\ref{tab:design}.
We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior; for example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
Several of the model parameters require special care and are reparametrized accordingly:
\begin{itemize}[leftmargin=1\parindent]
  \item The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eqn.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
    The observed nucleon fluctuation variance falls like $1 / n_c$ which means the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
    We therefore reparametrize the constituent fluctuations using the standard deviation of the resulting nucleon fluctuations
    \begin{equation}
      \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
    \end{equation}
    where $k$ is the shape parameter of the gamma distribution used to fluctuate each constituent.
  \item In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
    \begin{enumerate*}[label=(\roman*)]
      \item a parameter $v$ for the Gaussian width of each constituent,
      \item a parameter $r$ for the Gaussian width of the radial distribution used to sample the constituent centers, and
      \item a parameter $n_c$ to vary the number of constituents inside the nucleon.
    \end{enumerate*}
    In the limit $n_c \to \infty$, the composite nucleon's root-mean-square (RMS) radius is simply the convolution of its sampling radius $r$ and constituent width $v$ which add together in quadrature
  \begin{equation}
    \lim_{n_c \to\infty} r_\mathrm{RMS}\{\rho\} = \sqrt{r^2 + v^2}.
  \end{equation}
  We therefore choose to reparametrize the sampling radius $r$ in terms of a new variable
  \begin{equation}
    \label{eq:nucleon_width}
    w = \sqrt{r^2 + v^2},
  \end{equation}
  which approximates the RMS radius of the nucleon when the number of sampled constituents is large (see Fig.~\ref{fig:nucleon_schematic} for an example).
  We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} RMS radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior RMS radius.

@@

\item Equation~\eqref{eq:nucleon_width} requires the nucleon width to be larger than the constituent width, i.e. $w > v$, lest the sampling radius $r$ turn imaginary, and thus we cannot vary the nucleon width $w$ and the constituent width $v$ independently.
  We therefore reparametrize the constituent width $v$ using a new variable $\X$ to interpolate between minimum and maximum allowed values:
  \begin{equation}
    v = v_\mathrm{min} + \X (v_\mathrm{max} - v_\mathrm{min}).
  \end{equation}
  Here we choose a minimum width $v_\mathrm{min} = 0.2$~fm determined by computational limits and a maximum width $v_\mathrm{max} = w$ equal to the nucleon width.
  Thus for $\X=0$, the nucleons consist of small, distinct hot spots, whereas for $\X=1$ the nucleon is a single Gaussian blob of width $w$.
  The parameter $\X$ thus varies the granularity of the nucleon while keeping the number of constituents $n_c$ and approximate size of the nucleon $w$ fixed.
\end{itemize}

@@

However, for some observables such as $p$-Pb mean $p_T$ \cite{Abelev:2013bla} and flow cumulants $\vnk{n}{k}$ \cite{Chatrchyan:2013nka}, the experiments use a special high-multiplicity trigger to select rare, ultra-central events according to the number of charged particles produced or detector tracks offline.
These high-multiplicity bins are too selective for our modest minimum-bias event sample, and so a different procedure is required.
We exploit, for this purpose, the approximate correspondence
\begin{equation}
  \label{eq:bin_equiv}
  \ntrk \sim \nch \propto N_\rho
\end{equation}
between each event's initial secondary-parton multiplicity ${N_\rho = \int d^2x\,n(\x)}$ in Eqn.~\eqref{energy-momentum} and its final-state event activity, characterized by $\nch$ or $\ntrk$ \cite{Abelev:2013bla, Chatrchyan:2013nka}.

@@

We show simulated and emulated model observables (thin colored lines) for Pb-Pb collisions (Fig.~\ref{fig:obs_pbpb}) and \mbox{$p$-Pb} collisions (Fig.~\ref{fig:obs_ppb}) at $\sqrts=5.02$~TeV compared to experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
The top row of each figure shows explicit model calculations at each of the $d=500$ design points (training data), while the bottom row shows emulator predictions for $n=100$ random parameter samples drawn from the Bayesian posterior.
Each column shows a different class of observable.
The charged-particle yield $d\nch/d\eta$ is shown on the left, mean $p_T$ is in the middle, and two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$ are on the right.
The Pb-Pb mean $p_T$ and $p$-Pb $\vnk{4}{2}$ datasets are missing and hence are omitted from the present calibration.

@@

Conversely, the calibrated emulator predictions (bottom row of each figure) are far better constrained and nicely track the experimental data points.
We emphasize here that the posterior parameter values are obtained from a \emph{simultaneous} calibration to $p$-Pb and Pb-Pb data, and thus they are self-consistent between the two systems.
The spread in the posterior samples reflects different sources of model and experimental uncertainty as well as tension in the optimal fit parameters which describe each observable.
We demonstrate later in the text that a single set of model parameters well describes all of the calibration data, and thus we believe that much of the spread in the posterior samples is uncertainty contributed by our emulator.
We also note that although the $p$-Pb posterior samples have a somewhat larger spread than the Pb-Pb samples, the percentage uncertainty of the $p$-Pb emulator is similar to that of the Pb-Pb emulator, and thus the difference is likely due to the larger variance of the $p$-Pb training data.
The uncertainty in the posterior distribution could thus be improved by running the calibration with more design points or with a narrower range of parameter values to increase the density of the training points and reduce interpolation uncertainty.

We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
Recall that the posterior $P(H_\x | E)$ is the probability that our hypothesis $H_\x$ for the true model parameters $\x_\star$ is correct, given the evidence $E$ provided by experiment.
The present posterior has 15 dimensions, one for each parameter listed in Tbl.~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
We thus sample the distribution and histogram the samples to project the distribution onto one or two dimensions.
Each diagonal panel in Fig.~\ref{fig:posterior} is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
The posterior includes a wealth of information.
Here we focus on its most important features, starting with the top-left corner and working our way to the bottom-right.

@@

The \trento\ normalization factor is well constrained ${\text{Norm} = 20.3^{+2.2}_{-2.2}}$, as is the generalized mean parameter $p=0.018^{+0.146}_{-0.159}$ which fixes the scaling of initial QGP entropy deposition.

are well constrained, and we see that the posterior emulator predictions in Fig.~\ref{fig:obs_ppb} and \ref{fig:obs_pbpb} nicely describe the charged particle yields $d\nch/d\eta$ of both collision systems.
This is perhaps one of the strongest tests of the initial condition model.

which would not be the case if there was significant tension in fitting the $p$-Pb and Pb-Pb yields, or.
Moreover, our
It the emulator predictions using the posterior value describes both the $p$-Pb and Pb-Pb yields simultaneously (e.g.\ see Figs.~\ref{fig:obs_ppb}).
    We emphasize that this self-consistency is highly non-trivial since the ratio of $p$-Pb to Pb-Pb yields model can only fit both yields simultaneously if the \trento\ entropy deposition parameter $p$ is appropriately tuned.

>>>>

\newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
\newcommand{\sqrts}{\sqrt{s_{NN}}}
\newcommand{\nch}{N_\text{ch}}
\newcommand{\ntrk}{N_\text{trk}^\text{offline}}
\newcommand{\sigmaf}{\sigma_\mathrm{fluct}}
\newcommand{\X}{\chi_\mathrm{struct}}
\newcommand{\taufs}{\tau_\mathrm{fs}}
\newcommand{\dmin}{d_\mathrm{min}}
\newcommand{\T}{\tilde{T}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\vnk}[2]{v_#1\{#2\}}
\newcommand{\paddedhline}{\noalign{\smallskip}\hline\noalign{\smallskip}}
\newcommand{\order}[1]{$\mathcal O(10^{#1})$}
\newcommand{\note}{\textcolor{theblue}{[?]}}

@@

\begin{abstract}
We posit a unified hydrodynamic and microscopic description of the quark-gluon plasma (QGP) produced in ultrarelativistic $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference. Specifically, we model the dynamics of both collision systems using initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$ and flow cumulants.
We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that the flow produced develops at length scales smaller than a single proton.
Posterior estimates for the model's input parameters are obtained, and new insights into the temperature dependence of the QGP transport coefficients and event-by-event structure of the proton are discussed.
\end{abstract}

@@

Ultrarelativistic nuclear collisions between one light-ion and one heavy-ion, e.g.\ $^3$He-Au and $p$-Pb collisions, generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions where collectivity is commonly explained by the existence of hydrodynamic flow \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla, Adare:2015ctn}.
This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP) \cite{Bozek:2011if, Bozek:2013uha}, and that flow might develop at length scales smaller than a single proton \note.

@@

Independently, studies of coherent and incoherent $J/\psi$ production based on a color dipole picture of vector meson production were used to simultaneously constrain both the average color charge density of the proton as well as its event-by-event fluctuations in a saturation based framework \cite{Mantysaari:2016ykx, Mantysaari:2016jaz, Aaron:2009aa, Abramowicz:2015mha}.
Initial condition studies using the IP-Glasma model of color-glass condensate effective field theory \cite{Schenke:2012wb} simultaneously demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
In addition, it was recently shown that hydrodynamic simulations using IP-Glasma initial conditions with color charge fluctuations calibrated to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha} provide a good description of collectivity in small and large collision systems \cite{Schenke:2018qmc}.

@@

With these considerations in mind, we present progress towards a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free-streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$ and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.

@@

Assuming our two protons collide at the sampled impact parameter $\mathbf{b}$, we assign each a \emph{fluctuated} thickness
\begin{equation}
  \label{eq:fluctuated_thick}
  \T_{A, B}(\x) \equiv \int dz\, \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c \,(\mathbf{x} - \mathbf{x_i} \pm \mathbf{b}/2),
\end{equation}
equal to the beam-integrated proton density in Eqn.~\eqref{eq:rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.

Within the eikonal approximation, the initial entropy deposited at midrapidity at proper time $\tau=0^+$~fm/$c$ can be represented by some mapping ${f(\T_A, \T_B) \mapsto dS/dy}$.
A natural first guess for the mapping is the arithmetic mean
\begin{equation}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto \frac{\T_A + \T_B}{2}
\end{equation}
of the participant nuclear matter $\T_A$ and $T_B$, which is simply a wounded nucleon model up to meaningless factor of two in the normalization.
The wounded nucleon model was in fact one of the first such mappings used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thicknesses $\T_A$ and $\T_B$ \cite{Kharzeev:2000ph}.

A simple remedy is to replace the arithmetic mean of the wounded nucleon model with a more flexible parametrization
\begin{equation}
  \label{eq:gmean}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto \left( \frac{\T_A^p + \T_B^p}{2} \right)^{1/p},
\end{equation}
based on a family of functions known as the generalized mean(s).
This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric and harmonic means:
\begin{equation}
  \newlength{\extraspace}
  \setlength{\extraspace}{0.5ex}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto
  \begin{cases}
    \max(T_A, T_B) & p \rightarrow +\infty, \\[\extraspace]
    (T_A + T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
    \sqrt{T_A T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
    2\, T_A T_B/(T_A + T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
    \min(T_A, T_B) & p \rightarrow -\infty.
  \end{cases}
  \label{eq:gmean}
\end{equation}
Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
Of course the model is \emph{not} a substitute for first principle theory calculations, and it may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.

@@

\begin{figure}[b]
  \includegraphics{coupling}
  \caption{Cartoon of the free-streaming approximation for hydrodynamic initialization. The initial state is free-streamed for proper time $\taufs$ (zero coupling) before it is matched to hydrodynamics (strong coupling). This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
  \label{fig:coupling}
\end{figure}

@@

For the prior $P(H_\x)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Tbl.~\ref{tab:design}.
We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior; for example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
Several of the model parameters require special care and are reparametrized accordingly:
\begin{itemize}[leftmargin=1\parindent]
  \item The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eqn.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
    The observed nucleon fluctuation variance falls like $1 / n_c$ which means the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
    We therefore reparametrize the constituent fluctuations using the standard deviation of the resulting nucleon fluctuations
    \begin{equation}
      \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
    \end{equation}
    where $k$ is the shape parameter of the gamma distribution used to fluctuate each individual constituent.
  \item In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
    \begin{enumerate*}[label=(\roman*)]
      \item a parameter $v$ for the Gaussian width of each constituent,
      \item a parameter $r$ for the Gaussian width of the radial distribution used to sample the constituent centers, and
      \item a parameter $n_c$ to vary the number of constituents inside the nucleon.
    \end{enumerate*}
    In the limit $n_c \to \infty$, the composite nucleon's root-mean-square (RMS) radius is simply the convolution of its sampling radius $r$ and constituent width $v$ which add together in quadrature
  \begin{equation}
    \lim_{n_c \to\infty} r_\mathrm{RMS}\{\rho\} = \sqrt{r^2 + v^2}.
  \end{equation}
  We therefore choose to reparametrize the sampling radius $r$ in terms of a new variable
  \begin{equation}
    \label{eq:nucleon_width}
    w = \sqrt{r^2 + v^2},
  \end{equation}
  which approximates the RMS radius of the nucleon when the number of sampled constituents is large (see Fig.~\ref{fig:nucleon_schematic} for an example).
  We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} RMS radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior RMS radius.

@@

\item Equation~\eqref{eq:nucleon_width} requires the nucleon width to be larger than the constituent width, i.e. $w > v$, lest the sampling radius $r$ turn imaginary, and thus we cannot vary the nucleon width $w$ and the constituent width $v$ independently.
  We therefore reparametrize the constituent width $v$ using a new variable $\X$ to interpolate between minimum and maximum allowed values:
  \begin{equation}
    v = v_\mathrm{min} + \X (v_\mathrm{max} - v_\mathrm{min}).
  \end{equation}
  Here we choose a minimum constituent width $v_\mathrm{min} = 0.2$~fm determined by computational limits and a maximum width $v_\mathrm{max} = w$ equal to the nucleon width.
  Thus for $\X=0$, the nucleons consist of small, distinct hot spots, whereas for $\X=1$ the nucleon is a single Gaussian blob of width $w$.
  The parameter $\X$ thus varies the granularity of the nucleon while keeping the number of constituents $n_c$ and approximate size of the nucleon $w$ fixed.
\end{itemize}

@@

However, for some observables such as $p$-Pb mean $p_T$ \cite{Abelev:2013bla} and flow cumulants $\vnk{n}{k}$ \cite{Chatrchyan:2013nka}, the experiments use a special high-multiplicity trigger to select rare, ultra-central events according to the number of charged particles produced or detector tracks offline.
These high-multiplicity bins are too selective for our modest minimum-bias event sample, and so a different procedure is required.
We exploit, for this purpose, the approximate correspondence
\begin{equation}
  \label{eq:bin_equiv}
  \ntrk \sim \nch \propto N_\rho
\end{equation}
between each event's initial secondary-parton density ${N_\rho = \int d^2x\,n(\x)}$ in Eqn.~\eqref{energy-momentum} and its final-state event activity, characterized by $\nch$ or $\ntrk$ \cite{Abelev:2013bla, Chatrchyan:2013nka}.

@@

We show simulated and emulated model observables (thin colored lines) for Pb-Pb collisions (Fig.~\ref{fig:obs_pbpb}) and \mbox{$p$-Pb} collisions (Fig.~\ref{fig:obs_ppb}) at $\sqrts=5.02$~TeV compared to experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
The top row of each figure shows explicit model calculations at each of the $d=500$ design points (training data), while the bottom row shows emulator predictions for $n=100$ random parameter samples drawn from the Bayesian posterior (sampled from the MCMC chain).
Each column shows a different class of observable.
The charged-particle yield $d\nch/d\eta$ is shown on the left, mean $p_T$ is in the middle, and two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$ are on the right.
The Pb-Pb mean $p_T$ and $p$-Pb $\vnk{4}{2}$ datasets are missing and hence are omitted from the present calibration.

@@

Conversely, the calibrated emulator predictions (bottom row of Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}) are far better constrained and nicely track the experimental data points.
We emphasize here that the posterior parameter values are obtained from a \emph{simultaneous} calibration to $p$-Pb and Pb-Pb data, and thus they are self-consistent between the two systems.
The spread in the posterior samples reflects different sources of model and experimental uncertainty as well as tension in the optimal fit parameters which describe each observable.
We demonstrate later in the text that a single set of model parameters well describes all of the calibration data, and thus we believe that much of the spread in the posterior samples is uncertainty contributed by our emulator.
We also note that although the $p$-Pb posterior samples have a somewhat larger spread than the Pb-Pb samples, the percentage uncertainty of the $p$-Pb emulator is similar to that of the Pb-Pb emulator, and thus the difference is likely due to the larger variance of the $p$-Pb training data.
The uncertainty in the posterior distribution could thus be improved by running the calibration with more design points or with a narrower range of parameter values to increase the density of the training points and reduce interpolation uncertainty.

We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
Recall that the posterior $P(H_\x | E)$ is the probability that our hypothesis $H_\x$ for the true model parameters $\x_\star$ is correct, given the evidence $E$ provided by experiment.
The present posterior has 15 dimensions, one for each parameter listed in Tbl.~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
We thus sample the distribution and histogram the samples to project the distribution onto one or two dimensions at a time.
Each diagonal panel is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).

The posterior in Fig.~\ref{fig:posterior} includes a wealth of information.
Here we focus on its most important features, starting with the top-left corner of the figure and working our way to the bottom-right.
We report posterior estimates for individual parameters using their (marginalized) distribution medians and 90\% credible intervals.
For example, $x=2.45_{-0.15}^{+0.20}$ corresponds to a posterior median $\tilde{x}=2.45$ and 90\% credible interval $x \in (2.30, 2.65)$.

@@

The \trento\ normalization factor ${\text{Norm} = 20.3^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.018^{+0.146}_{-0.159}$ are well constrained by the present analysis.
Moreover, their posterior values nicely describe the calibration observables of both collision systems as evidenced by Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
While it would not be surprising, for example, to fit one or two of these observables using such a narrow range of values, the quality of the combined fit and the number of observables described is highly non-trivial.
For example, as the entropy deposition parameter $p \rightarrow \pm\infty$, particle production scales like the local maximum (minimum) of the nuclear overlap density.
This strongly affects the ratio of the predicted $p$-Pb and Pb-Pb yields and consequently---absent any additional information---excludes \emph{most} values of the parameter $p$.
It just so happens that the small range of $p$-values needed to describe the aforementioned yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \note.
This work thus reaffirms an empirical scaling law reported in several previous studies \note, that entropy deposition at midrapidity in ultrarelativistic nuclear collisions scales approximately as
\begin{equation}
  \label{eq:geometric_mean}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto \sqrt{\T_A \T_B},
\end{equation}
where $\T_A$ and $\T_B$ are the participant thickness functions \eqref{eq:fluctuated_thick} of each nucleus.
We emphasize that this specific analytic form should not be interpreted too literally.
For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible region, but it is not equal to the geometric mean in Eqn.~\ref{eq:geometric_mean}.
This approximate form, nevertheless, has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus it appears to be consistent with first-principle theoretical expectations.

Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the nucleon width parameter $w=0.97_{-0.17}^{+0.19}$~fm.
We caution that this parameter $w$ is \emph{not} the RMS radius of our nucleons due to idiosyncrasies of the constituent sampling procedure.
We can, however, easily calculate the RMS radius for a specific nucleon width $w$, constituent width $v$ and constituent number $n_c$.
For example, the single highest posterior probability region of parameter space prefers a nucleon width $w=0.92$~fm, constituent width $v=0.43$~fm and constituent number $n_c=6$.
The corresponding RMS nucleon radius for these parameters is $R_n = 0.86$~fm, conspicuously close the proton's RMS electric charge radius $R_p = 0.879(8)$~fm \cite{Bernauer:2010wm}.
